\Rpackage{MSnID} Package for Handling MS/MS Identifications

### 2014-04-08 inst/doc/msnid_vignette.Rnw
wholesale commit

\subsection{Analysis of peptide sequences}
A particular properties of peptide sequences we are interested in are 1) irregular cleavages at the termini of the peptides and 2) missing cleavage site within the peptide sequences. The default regular expressions of valid and missed cleavage patterns correspond to trypsin.
<<>>=
# creates new column NumIrregCleavages
msnid <- assess_termini(msnid, validCleavagePattern="[RK]\\.[^P]")
# creates new column NumMissCleavages
msnid <- assess_missed_cleavages(msnid, missedCleavagePattern="[KR](?=[^P$])")
@
Now the object has two more columns, NumIrregCleavages and NumMissCleavages, evidently corresponding to the number of termini with irregular cleavages and number of missed cleavages within the peptide sequence.
<<label=missedCleavages, fig=TRUE, include=FALSE, width=9>>=
pepCleav <- unique(psms(msnid)[,c("NumMissCleavages", "isDecoy", "Peptide")])
pepCleav <- as.data.frame(table(pepCleav[,c("NumMissCleavages", "isDecoy")]))
library("ggplot2")
ggplot(pepCleav, aes(x=NumMissCleavages, y=Freq, fill=isDecoy)) + 
   geom_bar(stat='identity', position='dodge') + 
   ggtitle("Number of Missed Cleavages")
@
\begin{center}
\includegraphics[width=0.8\textwidth]{graphics-missedCleavages}
\end{center}
\pagebreak[0]
Peptide sequences can by accessed by directly using \Rcode{\$} operator. For example:
<<label=lengths, fig=TRUE, include=FALSE, width=9>>=
# counting number of cysteins per peptide sequence
msnid$NumCys <- sapply(lapply(strsplit(msnid$Peptide,''),'==','C'),sum)
# calculating peptide lengths
msnid$PepLength <- nchar(msnid$Peptide) - 4
pepLen <- unique(psms(msnid)[,c("PepLength", "isDecoy", "Peptide")])
ggplot(pepLen, aes(x=PepLength, fill=isDecoy)) + 
   geom_histogram(position='dodge', binwidth=3) +
   ggtitle("Distribution on of Peptide Lengths")
@
\begin{center}
\includegraphics[width=0.8\textwidth]{graphics-lengths}
\end{center}
\pagebreak[0]

>>>>

\subsection{Analysis of peptide sequences}
A particular properties of peptide sequences we are interested in are 1) irregular cleavages at the termini of the peptides and 2) missing cleavage site within the peptide sequences. The default regular expressions of valid and missed cleavage patterns correspond to trypsin.
<<>>=
# creates new column NumIrregCleavages
msnid <- assess_termini(msnid, validCleavagePattern="[KR]\\.[^P]")
# creates new column NumMissCleavages
msnid <- assess_missed_cleavages(msnid, missedCleavagePattern="[KR](?=[^P$])")
@
Now the object has two more columns, NumIrregCleavages and NumMissCleavages, evidently corresponding to the number of termini with irregular cleavages and number of missed cleavages within the peptide sequence.
<<label=missedCleavages, fig=TRUE, include=FALSE, width=9>>=
pepCleav <- unique(psms(msnid)[,c("NumMissCleavages", "isDecoy", "Peptide")])
pepCleav <- as.data.frame(table(pepCleav[,c("NumMissCleavages", "isDecoy")]))
library("ggplot2")
ggplot(pepCleav, aes(x=NumMissCleavages, y=Freq, fill=isDecoy)) + 
   geom_bar(stat='identity', position='dodge') + 
   ggtitle("Number of Missed Cleavages")
@
\begin{center}
\includegraphics[width=0.8\textwidth]{graphics-missedCleavages}
\end{center}
\pagebreak[0]
Peptide sequences can by accessed by directly using \Rcode{\$} operator. For example:
<<label=lengths, fig=TRUE, include=FALSE, width=9>>=
# counting number of cysteins per peptide sequence
msnid$NumCys <- sapply(lapply(strsplit(msnid$Peptide,''),'==','C'),sum)
# calculating peptide lengths
msnid$PepLength <- nchar(msnid$Peptide) - 4
pepLen <- unique(psms(msnid)[,c("PepLength", "isDecoy", "Peptide")])
ggplot(pepLen, aes(x=PepLength, fill=isDecoy)) + 
   geom_histogram(position='dodge', binwidth=3) +
   ggtitle("Distribution on of Peptide Lengths")
@
\begin{center}
\includegraphics[width=0.8\textwidth]{graphics-lengths}
\end{center}
\pagebreak[0]
